% ==================================================================
% Minimal Example Chapter using the Archetype-Invariant Template
% This file is self-contained on environments; safe to drop-in.
% ==================================================================

\chapter{Localized Heat Trace near Internal Hypersurfaces}
\label{chap:lht-internal}

% --- minimal theorem envs if preamble lacks them
\makeatletter
\@ifundefined{theorem}{%
  \newtheorem{theorem}{Theorem}[chapter]
  \newtheorem{lemma}[theorem]{Lemma}
  \newtheorem{corollary}[theorem]{Corollary}
  \newtheorem{definition}[theorem]{Definition}
}{}
\makeatother

% --- archetype micro-macros (no-ops if already defined)
\providecommand{\Orientation}{\par\noindent\textbf{Orientation.}\ }
\providecommand{\Objectives}{\par\noindent\textbf{Objectives.}\ }
\providecommand{\Invariants}{\par\noindent\textbf{Invariants/Assumptions.}\ }
\providecommand{\Statement}{\par\noindent\textbf{Statement.}\ }
\providecommand{\ErrorBounds}{\par\noindent\textbf{Error bounds.}\ }
\providecommand{\Optimality}{\par\noindent\textbf{Optimality.}\ }
\providecommand{\Audit}{\par\noindent\textbf{Audit/Dependencies.}\ }
\providecommand{\Closure}{\par\noindent\textbf{Closure.}\ }

% ---------------- Orientation Block ----------------
\section*{Orientation}
We work under (SA.1--SA.4) and the notation of \Cref{sec:definitions}.
The goal is to state and prove a localized heat trace expansion for
the Dirichlet Laplacian on $M\setminus \Gamma$ with Dirichlet condition on
$\partial M\cup\Gamma$, isolating the fracture term and quantifying the remainder.

% ---------------- Foundational Objects ----------------
\section{Foundational Objects}

\begin{definition}[Fracture-adapted partition of unity]\label{def:fa-pou}
\Orientation Neighborhoods: a tubular neighborhood of $\Gamma$ and its complement.
\Statement Choose $\{\chi_{\mathrm{near}},\chi_{\mathrm{far}}\}\in C^\infty(M)$ with
$\chi_{\mathrm{near}}^2+\chi_{\mathrm{far}}^2\equiv 1$, $\operatorname{supp}\chi_{\mathrm{near}}$
contained in a fixed tubular neighborhood of $\Gamma$, and $\operatorname{supp}\chi_{\mathrm{far}}$
away from $\Gamma$. 
\Audit Compatible with reach$(\Gamma)>0$ and the $C^2$ graph structure.
\Closure Used to localize the parametrix and split the trace.
\end{definition}

\begin{definition}[Localized parametrix blocks]\label{def:blocks}
\Orientation Near $\Gamma$ use the doubled boundary model; away from $\Gamma$ use the standard boundary parametrix.
\Statement Define $K_{\mathrm{near}}(t,x,y)$ and $K_{\mathrm{far}}(t,x,y)$ as the respective heat kernel parametrices associated to the model problems, with standard short-time symbols.
\Audit Symbol classes match global $C^2$ regularity; constants depend only on $(d,\mathrm{reach},\|\mathrm{II}_\Gamma\|_\infty)$.
\Closure These feed the trace decomposition below.
\end{definition}

% ---------------- Auxiliary Results ----------------
\section{Auxiliary Results}

\begin{lemma}[Trace decomposition]\label{lem:trace-split}
\Orientation Localization with \Cref{def:fa-pou,def:blocks}.
\Statement For $t\downarrow 0$,
\[
\operatorname{Tr}(e^{-tL_\Gamma})
=\operatorname{Tr}\big(\chi_{\mathrm{near}} e^{-tL_\Gamma}\chi_{\mathrm{near}}\big)
+\operatorname{Tr}\big(\chi_{\mathrm{far}} e^{-tL_\Gamma}\chi_{\mathrm{far}}\big)
+O(e^{-c/t}).
\]
\Audit Off-diagonal terms are exponentially small by finite propagation/smoothing.
\Closure Reduces the analysis to two model contributions.
\end{lemma}

\begin{proof}
Insert the partition of unity under the trace and estimate mixed terms via
Gaussian off-diagonal bounds; the tubular geometry plus reach$>0$ gives the
required separation at short times.
\end{proof}

% ---------------- Main Results ----------------
\section{Main Results}

\begin{theorem}[Localized heat trace with fracture term]\label{thm:lht}
\Orientation $(M,g)$ compact, $\partial M$ piecewise $C^2$, $\Gamma\subset M$ closed $C^2$ embedded hypersurface, transverse to $\partial M$, reach$(\Gamma)>0$.
\Objectives Isolate $t^{-(d-1)/2}$ term from $\Gamma$ and quantify the remainder.
\Statement As $t\downarrow 0$,
\[
\operatorname{Tr}(e^{-tL_\Gamma})
=(4\pi t)^{-d/2}\,\mathrm{vol}_d(M)
-\tfrac14(4\pi t)^{-(d-1)/2}\,\mathrm{vol}_{d-1}(\partial M)
-\tfrac14(4\pi t)^{-(d-1)/2}\,\mathrm{vol}_{d-1}(\Gamma)
+O\!\big(t^{-(d-2)/2}\big).
\]
\ErrorBounds The implicit constant depends polynomially on the dimensionless complexity $\kappa(\Gamma)$ and on uniform geometric bounds (curvature, injectivity radius, reach).
\Audit Uses \Cref{lem:trace-split} and the model parametrices; constants tracked via symbol seminorms.
\Closure Supplies the canonical $a_\Gamma$ coefficient and the chapter’s target expansion.
\end{theorem}

\begin{proof}
Apply \Cref{lem:trace-split}; the far contribution reproduces the classical boundary terms.
Near $\Gamma$, use the doubled-boundary model and reflection to obtain the universal
$-\tfrac14(4\pi t)^{-(d-1)/2}$ density along $\Gamma$; integrate it to get
$-\tfrac14(4\pi t)^{-(d-1)/2}\,\mathrm{vol}_{d-1}(\Gamma)$. Remainder follows
from symbol bounds and uniform tubular control (reach$>0$).
\end{proof}

% ---------------- Corollaries and Consequences ----------------
\section{Corollaries and Consequences}

\begin{corollary}[Isotropic litho-invariant]\label{cor:KL}
\Orientation From \Cref{thm:lht}.
\Statement The isoperimetric normalization
\[
K_L^{\mathrm{iso}}(M,\Gamma)
:=\frac{a_\Gamma}{a_0^{(d-1)/d}}
= -\tfrac14\,\frac{\mathrm{vol}_{d-1}(\Gamma)}{\mathrm{vol}_d(M)^{(d-1)/d}}
\]
is scale-invariant and dimensionless.
\Audit $a_0=(4\pi)^{-d/2}\mathrm{vol}_d(M)$; $a_\Gamma$ as in \Cref{thm:lht}.
\Closure Provides the canonical dimensionless invariant used globally.
\end{corollary}

% ---------------- Closing Invariant ----------------
\section*{Closing Invariant}
We established the localized heat trace with a universal fracture term and
dimensionless normalization. This bridges the foundational definitions to the
dynamical refinements in the next chapter and is harmonized with \Cref{sec:definitions}.
